### 2015-06-20 RefereeResponse.md
Little typos

Dear Editor,

We thank the referee for a very detailed report. We feel that his/her
comments have dramatically improved the clarity and readability of our
paper.

We have addressed all the points raised by the referee. We include in
this reply an answer to each comment. We have included the file main_diff.pdf
showing the differences in the body of the paper.

The Authors.

Conceptual:
==

1. **It seems that part of the motivation for using FA is that it is
    very well correlated with lambda_1,  at low lambda_1. FA is a
    rather interesting quantity in itself, differentiating haloes and
    voids from  filaments and walls, but it is not clear why it should
    be superior to lambda_1 for void finding. The authors should more
    clearly articulate why this might be, if there is a reason. The FA
    being "conceived  to quantify the anisotropy of the diffusivity of
    water molecules through cerebral tissue in nuclear  magnetic
    resonance imaging" should be elaborated on; this does not clarify
    the usefulness of the FA for cosmologists. Conceptually, I also
    wonder: do filaments and walls behave differently from each other
    in  FA, or do they both have FA ~ 1?**

      Response:   
      - Although lambda_1 is the defining property of voids, 
      the FA has a key advantage that makes it a better void tracer,
      namely that it is possible to define a boundary in regions with
      FA~1. Boudary setting becomes arbitrary in the case of using
      lambda_1 as a tracer because lambda_1 is not bound. 
      We make this clarification in the  3rd paragraph of subsec. 3.2
      of the new version. 
      - Now it is explicitly explained why the FA, as defined in the
      context of MRI, is useful to cosmologist  
	(6th paragraph of introduction and 1st paragraph of subsec 3.1).
      - We prepare a new Figure (Fig. 3) to shows that filaments and
      sheets have different behaviour in FA. Only sheets are
      associated with high values of FA~1.

2. **I did not understand the sentence at the end of Sect 2, "In this
    work we propose an optimal value of  the threshold based on the
    maximization of the fractional anisotropy field in the locations
    label[ed]  as filaments and walls," and this seems to be an
    important point. Is the threshold mentioned in this  sentence a
    threshold in lambda_1, or FA? Is this the 0.95 threshold in FA? It
    seems to be a reasonable choice, but I do not see a sense in which
    it is "optimal," except that it seems to be the boundary  between
    the FA being well-correlated with lambda_1 and not-so-well. 
    Also, if this sentence is about the     0.95   
    threshold, I do not see why it "does not enter into our
    computations," since the boundary of many    voids will depend on
    this threshold. It would be useful to know for what fraction of
    voids the limit    FA = 0.95 actually affects the boundary (in the
    case of the largest voids, I suppose).**

    Response:
    We have rewritten this paragraph. The threshold that is 
    irrrelevant to our computations is lambda_th, not FA=0.95.

3. **I do not see how the value of FA=0.65 (corresponding to lambda_1
= 0) is used "to remove ridges."**

    Response: 
    - We have improved our explanation about the ridge removing
    procedure in the 3rd paragraph of subsection 3.3.

4. **In Fig. 6, I only see a clear ridge in FA at r~r_eff in a few
    cases. the sentence in Sect. 5.5 "The  difference between the
    radius where the density ridge is reached ([r/]reff = 3) and the
    radius of the  FA ridge ([r/]reff = 1 ) ..." implies that I should
    see a ridge in FA at r~r_eff. Is this because the ridge is
    actually a 2D surface that is spherically averaged? It would be
    helpful to point this (or any  other reasons) out. Also, the
    statement "This makes the FA ridge a reasonable boundary for voids
    compared  to the traditional definition that puts the boundary at
    the density ridge" implies (perhaps  unintentionally) that the FA
    ridge is a superior marker of the edge of a void than a density
    ridge. It is  a physically different sort of ridge, and I agree
    that it is reasonable, but I do not see why it should  be
    superior.**

    Response:
    - We now talk about "maxima" instead of "ridge" in the FA
    profiles. We have also removed the references of the FA as a
    "superior" characterization.

5. **I am not convinced that the "tiny white bubbles located inside
sheets" are embedded in overdense regions.  Based just on Fig. 1, they
could be physical, but small, voids. We also no nothing of their
extent  perpendicular to the plane of Fig. 1; we could just be looking
of corners of much larger objects.**

    Response:
    - In the right panels of Fig. 1, the color of each void refers now
    the effective radius. The size could be inferred from the
    figure. We have checked by visual inspection some cases and small
    voids are indeed located inside sheets and filaments. This was
    already found and explained in Hoffman et al. 2012.

6. **2D histograms of (lambda_1, delta) and (lambda_1, FA) are shown
    in Fig. 2, and I would appreciate also  seeing a (FA, delta) 2D
    histogram, maybe with 1+delta on a log scale, which would help to
    clarify the  physical meaning of FA.**

    Response: 
    - We have included the new plot in Fig. 2. This shows that indeed
    the overdensity of zero is found at the highest Fractional Anisotropy.

More minor things:
==

7. **I'm confused a bit about the middle panels of Fig. 1. The title,
    "Visual impresion for lambda_th=..."  is strange. Aren't there
    several thresholds here? Also, what are the blue dots? Also
    "impresion" is  misspelled.**

    Response:
    - We have rewritten the caption. There are two different
    thresholds, one for each web finder. Blue dots were dark matter
    halos (FOF). We removed them for clarity. The typo ("impresion")
    is  also corrected.

8. **Sect 3.3, 2nd paragraph: I do not see why "the analysis of [y]our
    results" shows that a fixed Cartesian  mesh does not induce
    spurious results". It's true, that the results look reasonable,
    but I think it would  be better to state something about the level
    of discreteness noise, e.g. the number of zero-density cells,  or
    the minimum if there are none.**

    Response:
    - We have rephrased this in terms of the number of particles per
    cell in the lowest density regions.

9. **Sect. 5.4: The sentence "subcompensated voids have outflowing ve-
    locity profiles all the way up to the effective radius where the
    average radial density reaches the average value" is a bit
    confusing; this is not r_eff, right? If so, this statement seems
    to be an understatement (v_r >0 out to several r_eff,
    typically). If not, "effective" should be removed for clarity.**

    Response:
    - We removed "effective" for clarity.

10. **In Sect. 5.5, there are a few instances where I think you meant
"r/r_eff =" instead of "r_eff =".**

    Response:
    - Indeed. We have revised all the instances and replaced r_eff by r/r_eff
    wherever it was appropriate.

11. **All quantities are computed at the pixel scale, ~1 Mpc/h,
    correct? If the authors have any insights about  how the results
    would change if this scale were changed, it would be helpful to
    state them. A quantitative  discussion of how some results would
    change with a different scale would be great, but not necessary.**

    Response:
    - We now mention in the first paragraph of the Results section the
    expected results for larger smoothing scales. Namely, that the
    statistics for the cosmci web tend to be those of a random
    Gaussian field.

12. **Finally, about the necessity of color: Figs 4-6, as they
currently are, require color, but it would be  easy for the authors to
use different linestyles, as well, to make color unnecessary.**

    Response:
    - These figures have already different linewidths for each bin. We
    think it might be easier to have 6 different widths than 6
    different linestyles.